A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines

### 2019-07-29 latex/paper.tex
Updates in abstract

\title{A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines}

@@

\begin{abstract}
% Topic
Recent increases in data volume and improved accessibility have 
led neuroimaging into the Big Data era. 
% Problem
While existing neuroimaging pipeline frameworks typically do not attempt to
limit data transfers, Big Data frameworks remain seldom used.
% Relevance
In this paper, we compare two popular Big Data frameworks with Python APIs,
Apache Spark and Dask, for their runtime performance in processing neuroimaging
pipelines.
% Approach
To evaluate performance differences we used two synthetic, Incrementation
(map-only) and Histogram (map-reduce), pipelines and real neuroimaging map-reduce pipeline (BIDS App Example). For both our 
synthetic pipelines we varied the number of workers and block size using the 
76~GB BigBrain image. We additionally varied the number of iterations and task
duration for our Incrementation pipeline. 39~GB of anatomical images from the
CoRR were used for the BIDS App Example pipeline.
% Methods (one sentence)
All experiments were executed on 8 instances of Compute Canada's Arbutus cloud
with 8~vCPUs and 30 GB of RAM all connected to a 2~TB shared NFS. We evaluated
PySpark's RDD against Dask's Bag, Delayed and Futures.
% Key Impact
Our result show that despite the differences between the two frameworks, both
frameworks perform the comparably. However, Dask pipelines risk being limited by
Python's GIL depending on task type and cluster configuration. In all cases, 
the major limiting factor was data transfer. While either framework is suitable
for neuroimaging pipelines, more effort needs to be placed in reducing data 
transfer time.
\end{abstract}

\begin{IEEEkeywords}
Big Data, Dask, Spark, performance, neuroimaging
\end{IEEEkeywords}

>>>>

\title{A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines}

@@

\begin{abstract}

In the past few years, neuroimaging has entered the Big Data era due to the
joint increase in image resolution, data sharing, and study sizes. However,
no particular Big Data framework has emerged in this field, and several
alternatives remain available. We compare two popular Big Data frameworks
with Python APIs, Apache Spark and Dask, for their runtime performance in
processing neuroimaging pipelines. Our evaluation uses two synthetic
pipelines processing the 76~GB Big Brain image, and a real pipeline
processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
 various combinations of task durations, data sizes, and numbers of workers,
deployed on a 64-node compute cluster in Compute Canada's Arbutus
cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
Results show that despite slight differences between Spark and Dask, both
frameworks perform comparably. However, Dask pipelines risk being limited
by Python's GIL depending on task type and cluster configuration. In all
cases, the major limiting factor was data transfer. While either framework
is suitable for neuroimaging pipelines, more effort needs to be placed in
reducing data transfer time.
\end{abstract}

\begin{IEEEkeywords}
Dask, Apache Spark, performance, neuroimaging
\end{IEEEkeywords}
